Manuscript draft

### 2024-05-06 manuscript.rmd
figures and further exploration for manuscript

# Introduction

```{r include = FALSE}
library(tidyverse)
library(here)
library(targets)
library(patchwork)
library(ggeffects)
library(gtsummary)
theme_set(theme_classic())
tar_config_set(store = here::here('_targets'))
#source(here("scripts/04.0_mixedModels.R"))
tar_load(cleaned_r)
tar_load(joined0_r)
tar_load(downsampled_10min_forSocial_r)
tar_load(downsampled_10min_r)
tar_load(cc)
tar_load(linked)
tar_load(deg_mod)
tar_load(deg_z_mod)
tar_load(str_mod)
tar_load(str_z_mod)
tar_load(situcolors)
tar_load(seasoncolors)
tar_load(movement_corr)
tar_load(space_corr)
tar_load(roost_corr)

@@

#### Social network centrality is known to be important in animals.

The social structure of animal groups has consequences for disease transmission, information sharing, and population dynamics. At the population level, the density and modularity of social interaction networks can drive epidemic probability and duration [@sah2017; @silk2017], govern the flow of information about food resources \[REF\], and contribute to population persistence through Allee effects \[REF\]. Individuals differ in their encounter rates, number of contacts, and centrality in the population social network \[REF\]. For individual animals, which make up the social environment of the group. social centrality in different behavioral situations contributes to their likelihood of contracting disease, finding mates, or accessing resources \[REFS\]. For example, in networks with high levels of degree heterogeneity, the most connected individuals may function as disease superspreaders [@lloyd-smith2005]. For example, possums with higher betweenness and closeness in a social network had higher infection rates [@corner2003], and higher connectivity in lizards was associated with a greater parasite load [@godfrey2009]. In general, individuals that are more central in the social network both have a higher infection risk and transmit disease more quickly than those that are more peripheral [@romano2016; @balasubramaniam2019].

#### We observe that some individuals are more central than others in a social network. What causes this variation?

@@

##### Individual traits

@@

#### Individuals' spatial phenotypes are important determinants of their social centrality

@@

-   Note that explicit spatial context is also important, and we didn't consider it. (We should have! How do we justify not doing that?)

@@

#### Social interactions occur when individuals encounter each other, in various behavioral situations.

@@

#### We set out to uncover whether spatial phenotypes could explain observed variation in social centrality.

@@

##### Research Questions

@@

# Methods {data-link="methods"}

### Study system

@@

## Field methods

@@

## Data preparation

@@

#### Determining roost locations

@@

#### Further cleaning for movement analysis

@@

## Measuring spatial phenotype

@@

## Summarizing movement measures

```{r include = FALSE}
# INSERT VARIABLE CONTRIBUTION PLOTS HERE
```

@@

**Movement** was defined as the first principal component of a PCA containing per-vulture seasonal means of daily mean flight altitude, daily flight duration, daily maximum displacement, and daily flight tortuosity. It explained [PERCENT]% of the variation in the PCA. Individuals with high movement scores traveled farther and flew longer, with straighter daily trajectories overall. **Space use** was defined as the first principal component of a PCA containing log-transformed 95% AKDE home range sizes and the ratio of the 50% AKDE to the 95% AKDE. It explained [PERCENT]% of the variation in the PCA. Individuals low on space use had small overall home ranges with concentrated core areas (and some exploratory movements away from the core), while individuals high on space use had large home ranges consisting of space that they used frequently (i.e. the core area made up a larger proportion of the home range). **Roost diversification** was defined as the first principal component of a PCA containing a seasonal Shannon diversity index of roost sites used \[ref Shannon\], number of unique roost sites used, and the proportion of nights in a given season when a vulture switched to a new roost site (versus roosting at the same site as on the previous night). It explained [PERCENT]% of the variation in the PCA.

@@

## Construction of proximity-based social networks

@@

### Co-flight

@@

### Co-feeding

@@

### Co-roosting

@@

## Calculation of social network measures

@@

## Comparison to a reference model

@@

## Model fitting and selection

@@

# Results

## Social preference (deviations from random sociality)

@@

```{r message = FALSE, out.width = "100%"}
linked %>%
  select(seasonUnique, situ, z_deg, z_str) %>%
  ungroup() %>%
  pivot_longer(cols = contains("z_"), names_to = "measure", values_to = "z") %>%
  ggplot(aes(x = z, col = situ, group = interaction(seasonUnique, measure, situ)))+
  geom_density(aes(linetype = measure))+
  facet_wrap(~measure*situ, scales = "free")+
  geom_vline(aes(xintercept = 0), col = "red")+
  scale_color_manual(name = "Situation", values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
  theme_classic()+
  theme(axis.title.y = element_blank())+
  xlab("Z-score")
```

## Degree

Preferential degree (i.e. degree z-score) was significantly associated with space use and roost diversification, but not with movement. Vultures that used more space (i.e. those with larger home ranges and larger proportional core areas) had a co-flight degree that deviated more from random (i.e. more preferential degree) than those with smaller space use. By contrast, co-roosting degree deviated less, not more, from random for wide-ranging vultures. [FIGURE] (Co-roosting degree showed a similar trend to co-flight, though it was only marginally significant).

```{r fig.cap = "Degree z-scores", out.width = "100%"}
deg_z_situ_space_emt_P <- effplot(model = deg_z_mod, terms = c("space_use", "situ"), 
                            dataset = linked, pointX = "space_use", 
                            pointY = "z_deg", col = "situ", 
                            legendTitle = "Situation", values = situcolors, 
                            ylab = "Degree (z-score)", xlab = "Space use (log)")

deg_z_situ_space_P <- emtplot(model = deg_z_mod, specs = "situ", 
                                var = "space_use", values = situcolors, 
                                ylab = "Space use effect (Degree z-scores)", 
                                xlab = "Situation")

deg_z_situ_space_emt_P + deg_z_situ_space_P + plot_layout(widths = c(4, 1), guides = "collect") & theme_classic()
```

@@

```{r}
deg_z_roost_P <- effplot(model = deg_z_mod, terms = c("roost_div", "situ"), 
                       dataset = linked, pointX = "roost_div", 
                       pointY = "z_deg", col = "situ", 
                       legendTitle = "Situation", values = situcolors, 
                       ylab = "Degree (z-score)", xlab = "Roost diversification")
deg_z_roost_P & theme_classic()
```

## Strength

The relationships were very similar for strength. Preferential strength in the co-flight and co-feeding networks was positively associated with space use, with wider-ranging vultures' strength deviating the most from random. Meanwhile, the opposite was once again true for the co-roosting network. Vultures with smaller ranges deviated singificantly more from random than those with larger ranges.

# Discussion

>>>>

Introduction

```{r include = FALSE}
library(tidyverse)
library(here)
library(targets)
library(patchwork)
library(ggeffects)
library(gtsummary)
library(extrafont)
loadfonts(device = "win")
theme_set(theme_classic() + theme(text = element_text(family="Verdana")))
tar_config_set(store = here::here('_targets'))
#source(here("scripts/04.0_mixedModels.R"))
tar_load(cleaned_r)
tar_load(joined0_r)
tar_load(downsampled_10min_forSocial_r)
tar_load(downsampled_10min_r)
tar_load(cc)
tar_load(linked)
tar_load(deg_mod)
tar_load(deg_z_mod)
tar_load(str_mod)
tar_load(str_z_mod)
tar_load(situcolors)
tar_load(seasoncolors)
tar_load(movement_corr)
tar_load(space_corr)
tar_load(roost_corr)

@@

Social network centrality is known to be important in animals.

The social structure of animal groups \[a\]has consequences for disease transmission, information sharing, and population dynamics. At the population level, the density and modularity of social interaction networks can drive epidemic probability and duration [@sah2017; @silk2017], govern the flow of information about food resources \[REF\], and contribute to population persistence through Allee effects \[REF\]. Individuals differ in their encounter rates, number of contacts, and centrality in the population social network \[REF\]. For individual animals, which make up the social environment of the group. social centrality in different behavioral situations contributes to their likelihood of contracting disease, finding mates, or accessing resources \[REFS\]. For example, in networks with high levels of degree heterogeneity, the most connected individuals may function as disease superspreaders [@lloyd-smith2005]. For example, possums with higher betweenness and closeness in a social network had higher infection rates [@corner2003], and higher connectivity in lizards was associated with a greater parasite load [@godfrey2009]. In general, individuals that are more central in the social network both have a higher infection risk and transmit disease more quickly than those that are more peripheral [@romano2016; @balasubramaniam2019].

We observe that some individuals are more central than others in a social network. What causes this variation?

@@

Individual traits

@@

Individuals' spatial phenotypes are important determinants of their social centrality

@@

@@

Social interactions occur when individuals encounter each other, in various behavioral situations.

@@

We set out to uncover whether spatial phenotypes could explain observed variation in social centrality.

@@

Research Questions

@@

Methods {data-link="methods"}

Study system

@@

Field methods

@@

Data preparation

@@

Determining roost locations

@@

Further cleaning for movement analysis

@@

Measuring spatial phenotype

@@

Summarizing movement measures

```{r eval = FALSE}
 INSERT VARIABLE CONTRIBUTION PLOTS HERE
```

@@

Movement was defined as the first principal component of a PCA containing per-vulture seasonal means of daily mean flight altitude, daily flight duration, daily maximum displacement, and daily flight tortuosity. It explained \[PERCENT\]% of the variation in the PCA. Individuals with high movement scores traveled farther and flew longer, with straighter daily trajectories overall. Space use was defined as the first principal component of a PCA containing log-transformed 95% AKDE home range sizes and the ratio of the 50% AKDE to the 95% AKDE. It explained \[PERCENT\]% of the variation in the PCA. Individuals low on space use had small overall home ranges with concentrated core areas (and some exploratory movements away from the core), while individuals high on space use had large home ranges consisting of space that they used frequently (i.e. the core area made up a larger proportion of the home range). Roost diversification was defined as the first principal component of a PCA containing a seasonal Shannon diversity index of roost sites used \[ref Shannon\], number of unique roost sites used, and the proportion of nights in a given season when a vulture switched to a new roost site (versus roosting at the same site as on the previous night). It explained \[PERCENT\]% of the variation in the PCA.

@@

Construction of proximity-based social networks

@@

Co-flight

@@

Co-feeding

@@

Co-roosting

@@

Calculation of social network measures

@@

Comparison to a reference model

@@

Model fitting and selection

@@

Results

Social preference (deviations from random sociality)

@@

```{r message = FALSE, out.width = "100%"}
linked %>%
  select(seasonUnique, situ, z_deg, z_str) %>%
  ungroup() %>%
  pivot_longer(cols = contains("z_"), names_to = "measure", values_to = "z") %>%
  ggplot(aes(x = z, col = situ, group = interaction(seasonUnique, measure, situ)))+
  geom_density(aes(linetype = measure))+
  facet_wrap(~measure*situ, scales = "free")+
  geom_vline(aes(xintercept = 0), col = "red")+
  scale_color_manual(name = "Situation", values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
  theme(axis.title.y = element_blank())+
  xlab("Z-score")
```

Degree

Preferential degree (i.e. degree z-score) was significantly associated with space use and roost diversification, but not with movement. Vultures that used more space (i.e. those with larger home ranges and larger proportional core areas) had a co-flight degree that deviated more from random (i.e. more preferential degree) than those with smaller space use. By contrast, co-roosting degree deviated less, not more, from random for wide-ranging vultures. \[FIGURE\] (Co-roosting degree showed a similar trend to co-flight, though it was only marginally significant).

```{r fig.cap = "Degree z-scores", out.width = "100%"}
# deg_z_situ_space_emt_P <- effplot(model = deg_z_mod, terms = c("space_use", "situ"), 
#                             dataset = linked, pointX = "space_use", 
#                             pointY = "z_deg", col = "situ", 
#                             legendTitle = "Situation", values = situcolors, 
#                             ylab = "Degree (z-score)", xlab = "Space use (log)")
# 
# deg_z_situ_space_P <- emtplot(model = deg_z_mod, specs = "situ", 
#                                 var = "space_use", values = situcolors, 
#                                 ylab = "Space use effect (Degree z-scores)", 
#                                 xlab = "Situation")
# 
# deg_z_situ_space_emt_P + deg_z_situ_space_P + plot_layout(widths = c(4, 1), guides = "collect") & theme_classic()
pts <- linked %>% mutate(group = case_when(situ == "Fe" ~ "Feeding",
                                           situ == "Fl" ~ "Flight",
                                           situ == "Ro" ~ "Roosting"))

dataset <- as.data.frame(ggeffect(model = deg_z_mod, terms = c("space_use", "situ"))) %>% mutate(group = case_when(group == "Fe" ~ "Feeding",
                                                                                                                   group == "Fl" ~ "Flight",
                                                                                                                   group == "Ro" ~ "Roosting"))

a <- ggplot(data = dataset, 
       aes(x, predicted))+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
              alpha = 0.2, linewidth = 0.6, show.legend = F)+
  geom_point(data = pts,
             aes(x = space_use, y = z_deg, col = group),
             pch = 1)+
  geom_line(aes(col = group), linewidth = 1)+
  facet_wrap(~group, scales = "free_y")+
  scale_color_manual(name = "Situation", values = situcolors)+
  scale_fill_manual(name = "Situation", values = situcolors)+
  ylab("Degree (z-score)")+ xlab("Space use (log)")+
  theme(legend.position = "none",
        text = element_text(size = 16))

dat <- as.data.frame(emmeans::emtrends(deg_z_mod, specs = "situ", var = "space_use")) %>% mutate(situ = case_when(situ == "Fe" ~ "Feeding",
                                                                                                                  situ == "Fl" ~ "Flight",
                                                                                                                  situ == "Ro" ~ "Roosting"))

b <- dat %>%
  ggplot(aes(x = situ, y = space_use.trend, col = situ))+
  geom_point(size = 6)+
  geom_errorbar(aes(x = situ, ymin = lower.CL, ymax = upper.CL), 
                width = 0, linewidth = 2) +
  geom_hline(aes(yintercept = 0), linetype = 2)+
  scale_color_manual(values = situcolors)+ scale_x_discrete(limits = rev, position = "top")+
  coord_flip()+ ylab("Space use \neffect")+
  theme(legend.position = "none", 
        text = element_text(size = 16),
        axis.title.y = element_blank())

plt <- a + b + plot_layout(widths = c(4, 1))
ggsave(plot = plt, filename = here("fig/figures/degree/sp_situ.png"), width = 12, height = 6)
```

@@

```{r}
dataset <- as.data.frame(ggeffect(model = deg_z_mod, terms = c("roost_div", "situ"))) %>% mutate(group = case_when(group == "Fe" ~ "Feeding",
                                                                                                                   group == "Fl" ~ "Flight",
                                                                                                                   group == "Ro" ~ "Roosting"))

a <- ggplot(data = dataset, 
       aes(x, predicted))+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
              alpha = 0.2, linewidth = 0.6, show.legend = F)+
  geom_point(data = pts,
             aes(x = space_use, y = z_deg, col = group),
             pch = 1)+
  geom_line(aes(col = group), linewidth = 1)+
  facet_wrap(~group, scales = "free_y")+
  scale_color_manual(name = "Situation", values = situcolors)+
  scale_fill_manual(name = "Situation", values = situcolors)+
  ylab("Degree (z-score)")+ xlab("Roost diversification")+
  theme(legend.position = "none",
        text = element_text(size = 16))

dat <- as.data.frame(emmeans::emtrends(deg_z_mod, specs = "situ", var = "roost_div")) %>% mutate(situ = case_when(situ == "Fe" ~ "Feeding",
                                                                                                                  situ == "Fl" ~ "Flight",
                                                                                                                  situ == "Ro" ~ "Roosting"))

b <- dat %>%
  ggplot(aes(x = situ, y = roost_div.trend, col = situ))+
  geom_point(size = 6)+
  geom_errorbar(aes(x = situ, ymin = lower.CL, ymax = upper.CL), 
                width = 0, linewidth = 2) +
  geom_hline(aes(yintercept = 0), linetype = 2)+
  scale_color_manual(values = situcolors)+ scale_x_discrete(limits = rev, position = "top")+
  coord_flip()+ ylab("Roost diversification \neffect")+
  theme(legend.position = "none", 
        text = element_text(size = 16),
        axis.title.y = element_blank())

plt <- a + b + plot_layout(widths = c(4, 1))
ggsave(plot = plt, filename = here("fig/figures/degree/rd_situ.png"), width = 12, height = 6)
```

Strength

The relationships were very similar for strength. Preferential strength in the co-flight and co-feeding networks was positively associated with space use, with wider-ranging vultures' strength deviating the most from random. Meanwhile, the opposite was once again true for the co-roosting network. Vultures with smaller ranges deviated significantly more from random than those with larger ranges.

```{r}
dataset <- as.data.frame(ggeffect(model = str_z_mod, terms = c("space_use", "situ"))) %>% mutate(group = case_when(group == "Fe" ~ "Feeding",
                                                                                                                   group == "Fl" ~ "Flight",
                                                                                                                   group == "Ro" ~ "Roosting"))

a <- ggplot(data = dataset, 
       aes(x, predicted))+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
              alpha = 0.2, linewidth = 0.6, show.legend = F)+
  geom_point(data = pts,
             aes(x = space_use, y = z_str, col = group),
             pch = 1)+
  geom_line(aes(col = group), linewidth = 1)+
  facet_wrap(~group, scales = "free_y")+
  scale_color_manual(name = "Situation", values = situcolors)+
  scale_fill_manual(name = "Situation", values = situcolors)+
  ylab("Strength (z-score)")+ xlab("Space use (log)")+
  theme(legend.position = "none",
        text = element_text(size = 16))

dat <- as.data.frame(emmeans::emtrends(str_z_mod, specs = "situ", var = "space_use")) %>% mutate(situ = case_when(situ == "Fe" ~ "Feeding",
                                                                                                                  situ == "Fl" ~ "Flight",
                                                                                                                  situ == "Ro" ~ "Roosting"))

b <- dat %>%
  ggplot(aes(x = situ, y = space_use.trend, col = situ))+
  geom_point(size = 6)+
  geom_errorbar(aes(x = situ, ymin = lower.CL, ymax = upper.CL), 
                width = 0, linewidth = 2) +
  geom_hline(aes(yintercept = 0), linetype = 2)+
  scale_color_manual(values = situcolors)+ scale_x_discrete(limits = rev, position = "top")+
  coord_flip()+ ylab("Space use \neffect")+
  theme(legend.position = "none", 
        text = element_text(size = 16),
        axis.title.y = element_blank())

plt <- a + b + plot_layout(widths = c(4, 1))
ggsave(plot = plt, filename = here("fig/figures/strength/sp_situ.png"), width = 12, height = 6)
```

Similarly, roost diversification negatively predicted preferential strength \[...\]

```{r}
dataset <- as.data.frame(ggeffect(model = str_z_mod, terms = c("roost_div", "situ"))) %>% mutate(group = case_when(group == "Fe" ~ "Feeding",
                                                                                                                   group == "Fl" ~ "Flight",
                                                                                                                   group == "Ro" ~ "Roosting"))

a <- ggplot(data = dataset, 
       aes(x, predicted))+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
              alpha = 0.2, linewidth = 0.6, show.legend = F)+
  geom_point(data = pts,
             aes(x = space_use, y = z_str, col = group),
             pch = 1)+
  geom_line(aes(col = group), linewidth = 1)+
  facet_wrap(~group, scales = "free_y")+
  scale_color_manual(name = "Situation", values = situcolors)+
  scale_fill_manual(name = "Situation", values = situcolors)+
  ylab("Strength (z-score)")+ xlab("Roost diversification")+
  theme(legend.position = "none",
        text = element_text(size = 16))

dat <- as.data.frame(emmeans::emtrends(str_z_mod, specs = "situ", var = "roost_div")) %>% mutate(situ = case_when(situ == "Fe" ~ "Feeding",
                                                                                                                  situ == "Fl" ~ "Flight",
                                                                                                                  situ == "Ro" ~ "Roosting"))

b <- dat %>%
  ggplot(aes(x = situ, y = roost_div.trend, col = situ))+
  geom_point(size = 6)+
  geom_errorbar(aes(x = situ, ymin = lower.CL, ymax = upper.CL), 
                width = 0, linewidth = 2) +
  geom_hline(aes(yintercept = 0), linetype = 2)+
  scale_color_manual(values = situcolors)+ scale_x_discrete(limits = rev, position = "top")+
  coord_flip()+ ylab("Roost diversification \neffect")+
  theme(legend.position = "none", 
        text = element_text(size = 16),
        axis.title.y = element_blank())

plt <- a + b + plot_layout(widths = c(4, 1))
ggsave(plot = plt, filename = here("fig/figures/strength/rd_situ.png"), width = 12, height = 6)
```

## How much does z-score vary by individual across situations and seasons?

```{r}
ll <- linked %>%
  select(Nili_id, seasonUnique, season, year, situ, age_group, sex, z_deg, z_str)

pltfun <- function(sit){
  dat <- ll %>%
    filter(situ == sit) %>%
    group_by(Nili_id) %>%
    mutate(mn = mean(z_deg)) %>%
    ungroup()

  highlight <- dat %>% group_by(Nili_id) %>%
    filter(n() > 4) %>%
    ungroup() %>%
    filter(mn %in% c(min(mn), max(mn), median(mn))) %>%
    pull(Nili_id) %>% unique()

  dat <- dat %>%
    mutate(hl = ifelse(Nili_id %in% highlight, T, F))

  plt <- dat %>%
    ggplot(aes(x = seasonUnique, y = z_deg, col = mn))+
    geom_jitter(pch = 1, width = 0.1, alpha = 0.2)+
    geom_line(aes(group = Nili_id, alpha = hl, linewidth = hl))+
    scale_color_viridis_c()+
    scale_alpha_manual(values = c(0.2, 1))+
    scale_linewidth_manual(values = c(0.2, 1))
  return(plt)
}

```

Looks like there might be a little bit of repeatability within an individual, but there's still a lot of variation as well.

What about correlation between z-score in one social situation and z-score in the others?

```{r}
llw <- ll %>%
  select(Nili_id, seasonUnique, situ, z_deg) %>%
  pivot_wider(id_cols = c("Nili_id", "seasonUnique"), names_from = "situ", values_from = "z_deg")

llw %>%
  ggplot(aes(x = Fe, y = Fl, col = seasonUnique))+geom_point(alpha = 0.5, pch = 1)+geom_smooth(method = "lm", se = F, linewidth = 0.5) # there is generally a positive correlation between feeding and flight

llw %>%
  ggplot(aes(x = Fl, y = Ro, col = seasonUnique))+geom_point(alpha = 0.5, pch = 1)+geom_smooth(method = "lm", se = F, linewidth = 0.5) # generally a negative correlation between flight and roosting (makes sense, given the opposite trends)

llw %>%
  ggplot(aes(x = Ro, y = Fe, col = seasonUnique))+geom_point(alpha = 0.5, pch = 1)+geom_smooth(method = "lm", se = F, linewidth = 0.5) # Also a bit of a negative relationship here.
```
